---
title: "Aarti_Evolution_Project"
author: "Truc Minh Nguyen"
date: "2025-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
library(tidyverse)
```

```{r}

adapt_form <- read.csv("adaptationdata_wdictionary.csv")
spec_form <- read.csv("speciationdata_wdictionary.csv")
  
```

## Whether pre-test science conceptions are stronger predictors of a final test grade than pre-test science identity or beliefs about the fixedness of intelligence (fixed mindset)

## Start with looking at post-scores vs. pre-test science conceptions
```{r}
# Post-scores vs. Pre-science Conceptions - Evidence for Adaptation
ggplot(adapt_form, aes(x = factor(scievidence_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Conceptions - Evidence Adaption Curriculum",
    x = "Pre-Science Evidence Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science evidence ratings
table(adapt_form$scievidence_pre)
```

```{r}
# Post-scores vs. Pre-science Conceptions - Evidence for Speciation
ggplot(spec_form, aes(x = factor(scievidence_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Conceptions - Evidence Speciation Curriculum",
    x = "Pre-Science Evidence Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science evidence ratings
table(spec_form$scievidence_pre)
```


```{r}
# Post-scores vs. Pre-science Conceptions - Tools for Adaptation
ggplot(adapt_form, aes(x = factor(scitools_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Conceptions - Tools Adaptation Curriculum",
    x = "Pre-Science Tools Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science Tools ratings
table(adapt_form$scitools_pre)

```


```{r}
# Post-scores vs. Pre-science Conceptions - Tools for Speciation
ggplot(spec_form, aes(x = factor(scitools_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Conceptions - Tools Speciation Curriculum",
    x = "Pre-Science Tools Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science Tools ratings
table(spec_form$scitools_pre)

```



```{r}
# Post-scores vs. Pre-science Conceptions - Explain Adaptation
ggplot(adapt_form, aes(x = factor(sciexplain_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Conceptions - Explanation Adaptation Curriculum",
    x = "Pre-Science Explanation Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science Explanations ratings
table(adapt_form$sciexplain_pre)

```
Notes: I see that the counts for the pre-science conceptions are higher in the 
rating category of 3 and 4 rather than 1 and 2. So the majority appears to initially
agree with these science conception statements. The median for post scores appear to 
be close to each other regardless of the pre-ratings category but the spread in 
each category is a little different but the counts are also different..


```{r}

# Post-scores vs. Pre-science Conceptions - Explain Speciation
ggplot(spec_form, aes(x = factor(sciexplain_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Conceptions - Explanation Speciation Curriculum",
    x = "Pre-Science Explanation Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science Explanations ratings
table(spec_form$sciexplain_pre)

```


## Next, look at post-test scores vs. pre-science identity ratings.

```{r}
# Post-scores vs. Pre-science Identity - Self Adaptation 
ggplot(adapt_form, aes(x = factor(selfsciid_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Identity - Self Adaptation Curriculum",
    x = "Pre-Science Self-Identity Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science Identity-self ratings
table(adapt_form$selfsciid_pre)

```


```{r}
# Post-scores vs. Pre-science Identity - Self Speciation
ggplot(spec_form, aes(x = factor(selfsciid_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Identity - Self Speciation Curriculum",
    x = "Pre-Science Self-Identity Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science Identity-self ratings
table(spec_form$selfsciid_pre)
```


```{r}
# Post-scores vs. Pre-science Identity - Others Adaptation
ggplot(adapt_form, aes(x = factor(othersciid_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Identity - Others Adaptation Curriculum",
    x = "Pre-Science Others-Identity Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science identity-others ratings
table(adapt_form$othersciid_pre)

```

Notes: Interesting that they view themselves higher in science than what they 
think others view them in science. Because the count for others rating in 
category 4 is 18 for others vs. 83 for self.. 


```{r}

# Post-scores vs. Pre-science Identity - Others Speciation
ggplot(spec_form, aes(x = factor(othersciid_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-science Identity - Others Speciation Curriculum",
    x = "Pre-Science Others-Identity Ratings",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science identity-others ratings
table(spec_form$othersciid_pre)

```


## Let's look at post-scores and pre-mindset composite scores
```{r}
# Post-scores vs. Pre-mindset composite scores Adaptation
ggplot(adapt_form, aes(x = factor(mindset_comp_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-Mindset Composite Scores Adaptation Curriculum",
    x = "Pre-Mindset Composite Scores",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science identity-others ratings
table(adapt_form$mindset_comp_pre)

```

Notes: the outliers in category 1.33, 1.667, and 3 might be incomplete post test 
where there were no-response..

```{r}
# Post-scores vs. Pre-mindset composite scores Speciation
ggplot(spec_form, aes(x = factor(mindset_comp_pre), y = percentage_post)) +
  geom_boxplot(fill = "skyblue", color = "darkblue") +
  labs(
    title = "Post-scores vs. Pre-Mindset Composite Scores Speciation Curriculum",
    x = "Pre-Mindset Composite Scores",
    y = "Post-scores"
  ) +
  theme_minimal()

# Create table showing observation counts of Pre-science identity-others ratings
table(spec_form$mindset_comp_pre)

```


### Let's look at the correlations among predictors and response variable

```{r}

# Subset the data to include only response and selected independent variables
cor_data <- adapt_form %>%
  select(percentage_post, scievidence_pre, scitools_pre, sciexplain_pre,
         selfsciid_pre, othersciid_pre,
         mindset_comp_pre)  


cor_matrix <- cor(cor_data, use = "complete.obs", method = "pearson")

print(cor_matrix)

# Visualize correlation matrix
corrplot(cor_matrix, method = "color", 
         type = "upper",          # Show only upper triangle
         tl.col = "black",        # Label color
         tl.srt = 45,             # Rotate label text
         addCoef.col = "black")   # Add correlation coefficients


```

Notes: doesn't seems alike any of these predictors have strong correlation with 
response variable percentage_post. But some predictors seem to have strong correlation 
with each other such as pre-science conception_evidence and pre-science 
conception_explanation (0.45), and pre-self identity and pre-other identity (0.52).

### Let's model with original dataset first 

```{r}
model_base_adapt <- lm(percentage_post ~ factor(selfsciid_pre) + factor(othersciid_pre) + 
                   mindset_comp_pre, data = adapt_form)
summary(model_base_adapt)

model_full_adapt <- lm(percentage_post ~ factor(selfsciid_pre) + factor(othersciid_pre) 
                 + mindset_comp_pre + factor(scievidence_pre) + 
                   factor(scitools_pre) + factor(sciexplain_pre), 
                 data = adapt_form)
summary(model_full_adapt)
anova(model_base_adapt,model_full_adapt)

```



```{r}
model_base_spec <- lm(percentage_post ~ factor(selfsciid_pre) + factor(othersciid_pre) + 
                   mindset_comp_pre, data = spec_form)
summary(model_base_spec)

model_full_spec <- lm(percentage_post ~ factor(selfsciid_pre) + factor(othersciid_pre) 
                 + mindset_comp_pre + factor(scievidence_pre) + 
                   factor(scitools_pre) + factor(sciexplain_pre), 
                 data = spec_form)
summary(model_full_spec)
anova(model_base_spec,model_full_spec)

```


### Let's pivot data long to for modeling. 
```{r}
adapt_long <- adapt_form %>%
  pivot_longer(
    cols = -c(ID, SCHOOL, TEACHER, SES,curriculum_type), # Keep static columns
    names_to = c("Variable", "Time_Pre_Post"), # Create new columns
    names_pattern = "(.+)_(pre|post)" # Extract variable name and time (pre/post)
  )

# Reshape the long format data to wide format by spreading Variable into columns
adap_long_final <- adapt_long %>%
  pivot_wider(
    names_from = Variable,  # Create new columns for each unique Variable
    values_from = value     # Fill them with corresponding values
  )

```

