---
title: "Aarti0219"
author: "Haochen_Li"
date: "2025-02-19"
output: pdf_document
---

Reshaping the dataset

```{r}
library(tidyverse)
library(lme4)
library(car)
library(DHARMa)
library(MASS)
library(glmmTMB)
library(ordinal)
library(nnet)
library(lmtest)
library(caret)
library(cmdstanr)
library(rstanarm)
library(brms)
library(performance)
library(loo)
library(ggplot2)
library(bayesplot)
df_long$scievidence  <- factor(df_long$scievidence, ordered = FALSE)
df_long$scitools     <- factor(df_long$scitools, ordered = FALSE)
df_long$sciexplain   <- factor(df_long$sciexplain, ordered = FALSE)
df_long$ID <- as.factor(df_long$ID)
df_long$SCHOOL <- as.factor(df_long$SCHOOL)
```

```{r}
setwd("E:/Desktop/BU/2025 Spring/MA676/Consult")
df <- read.csv("merge_fmt.csv")

# Reshape
pre_post_columns <- names(df)[grepl("_pre|_post", names(df))]
percentage_columns <- c("percentage_pre", "percentage_post")

# Convert to long format while keeping original variable names
df_long <- df %>%
  pivot_longer(
    cols = setdiff(pre_post_columns, percentage_columns),  # Exclude percentage columns from reshaping
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Time = ifelse(grepl("_pre$", Variable), 0, 1),  # Assign 0 for pre, 1 for post
    Variable = gsub("_pre|_post", "", Variable)  # Remove _pre and _post to keep original variable names
  ) %>%
  pivot_wider(names_from = Variable, values_from = Value)  # Spread variables back into wide format

# Retain percentage_pre and percentage_post while reshaping
df_long <- df_long %>%
  group_by(ID) %>%  # Ensure values are assigned correctly per ID
  mutate(
    percentage_pre = unique(percentage_pre, na.rm = TRUE),
    percentage_post = unique(percentage_post, na.rm = TRUE)
  ) %>%
  ungroup()

df_long$scievidence  <- factor(df_long$scievidence, ordered = TRUE)
df_long$scitools     <- factor(df_long$scitools, ordered = TRUE)
df_long$sciexplain   <- factor(df_long$sciexplain, ordered = TRUE)
df_long$ID <- as.factor(df_long$ID)
df_long$SCHOOL <- as.factor(df_long$SCHOOL)
```
(2) Whether kids participation in a longer vs. shorter version of a challenging curriculum changes three general conceptions of science

```{r}
model_scitools <- ordinal::clmm(
  scitools ~ Time * session + (1 | ID) + (1 | SCHOOL), 
  data = df_long, 
  link = "logit",
  control = clmm.control(method = "nlminb")
)

model_scievidence <- ordinal::clmm(
  scievidence ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  link = "logit",
  control = clmm.control(method = "nlminb")
)
model_sciexplain <- ordinal::clmm(
  sciexplain ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  link = "logit",
  control = clmm.control(method = "nlminb")
)

summary(model_scitools)
summary(model_scievidence)
summary(model_sciexplain)
library(brant)
brant(model_scitools)

```

model_scitools: No significant effects were found, suggesting that students' conceptions of science tools did not change based on time (pre vs. post) or curriculum length.

model_scievidence: The length of the curriculum had a significant effect on students’ perceptions of scientific evidence (p = 0.0437). However, there was no evidence that students' perceptions changed over time (p = 0.1353).

model_sciexplain: No significant effects were found in this model. Students' perceptions of scientific explanations did not change over time, nor were they affected by the curriculum length.

```{r}
model_scitools_npo <- ordinal::clmm(
  scitools ~ Time * session + factor(scitools) * Time + factor(scitools) * session + (1 | ID) + (1 | SCHOOL), 
  data = df_long, 
  link = "logit",
  control = clmm.control(method = "nlminb")
)
model_scievidence_npo <- ordinal::clmm(
  scievidence ~ Time * session + factor(scievidence) * Time + factor(scievidence) * session + (1 | ID)+ (1 | SCHOOL), 
  data = df_long, 
  link = "logit",
  control = clmm.control(method = "nlminb")
)
model_sciexplain_npo <- ordinal::clmm(
  sciexplain ~ Time * session + factor(sciexplain) * Time + factor(sciexplain) * session+ (1 | ID)+ (1 | SCHOOL), 
  data = df_long, 
  link = "logit",
  control = clmm.control(method = "nlminb")
)
summary(model_scitools_npo)
summary(model_scievidence_npo)
summary(model_sciexplain_npo)
```


Distribution Plot
```{r}
df_long_3cpt <- df_long %>%
  pivot_longer(cols = c(scitools, scievidence, sciexplain), 
               names_to = "Variable", 
               values_to = "Value")


ggplot(df_long_3cpt, aes(x = as.factor(Value), fill = Variable)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ Variable, scales = "free_x") +
  labs(title = "Distribution of Scitools, Scievidence, and Sciexplain",
       x = "Category",
       y = "Count") +
  theme_minimal()

```


LRT test for scitools
```{r}
model_npo <- clmm(
  scitools ~ Time * session + factor(scitools) + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  link = "logit"
)
anova(model_scitools, model_npo)
VarCorr(model_scitools)
```
LRT test for scievidence
```{r}
model_npo <- clmm(
  scievidence ~ Time * session + factor(scievidence) + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  link = "logit"
)
anova(model_scitools, model_npo)
VarCorr(model_scitools)
```
LRT test for sciexplain
```{r}
model_npo <- clmm(
  sciexplain ~ Time * session + factor(sciexplain) + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  link = "logit"
)
anova(model_scitools, model_npo)
VarCorr(model_scitools)
```

Model comparison

```{r}
model_scitools_noSchool <- clmm(
  scitools ~ Time * session + (1 | ID),
  data = df_long,
  link = "logit"
)
AIC(model_scitools, model_scitools_noSchool)

model_scievidence_noSchool <- clmm(
  scitools ~ Time * session + (1 | ID),
  data = df_long,
  link = "logit"
)
AIC(model_scievidence, model_scievidence_noSchool)

model_sciexplain_noSchool <- clmm(
  scitools ~ Time * session + (1 | ID),
  data = df_long,
  link = "logit"
)
AIC(model_sciexplain, model_sciexplain_noSchool)
```




preprocess grouped

```{r}
df_long$scitools_grouped <- factor(ifelse(df_long$scitools %in% c(1, 2), "1_2", as.character(df_long$scitools)),
                                    levels = c("1_2", "3", "4"))
df_long$scievidence_grouped <- factor(ifelse(df_long$scievidence %in% c(1, 2), "1_2", as.character(df_long$scievidence)),
                                    levels = c("1_2", "3", "4"))
df_long$sciexplain_grouped <- factor(ifelse(df_long$sciexplain %in% c(1, 2), "1_2", as.character(df_long$sciexplain)),
                                    levels = c("1_2", "3", "4"))
df_long$scitools_grouped <- factor(df_long$scitools_grouped, ordered = TRUE)
df_long$scievidence_grouped <- factor(df_long$scievidence_grouped, ordered = TRUE)
df_long$sciexplain_grouped <- factor(df_long$sciexplain_grouped, ordered = TRUE)
```

Baysian approach

```{r}
scitools_grouped_model <- brm(
  scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit"),
  chains = 4, iter = 4000, warmup = 500, cores = 8,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

scitools_grouped_npo_model <- brm(
  scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  save_pars = save_pars(all = TRUE),
  chains = 4, iter = 4000, cores = 8
)

scievidence_grouped_model <- brm(
  scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit"),
  chains = 4, iter = 4000, warmup = 500, cores = 8,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

scievidence_grouped_npo_model <- brm(
  scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  save_pars = save_pars(all = TRUE),
  chains = 4, iter = 4000, cores = 8
)
sciexplain_grouped_model <- brm(
  sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit"),
  chains = 4, iter = 4000, warmup = 500, cores = 8,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

sciexplain_grouped_npo_model <- brm(
  sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  save_pars = save_pars(all = TRUE),
  chains = 4, iter = 4000, cores = 8
)
scitools_grouped_model <- brm(
  scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit"),
  chains = 4, iter = 4000, warmup = 500, cores = 8,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

scitools_grouped_npo_model <- brm(
  scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  save_pars = save_pars(all = TRUE),
  chains = 4, iter = 4000, cores = 8
)

scievidence_grouped_model <- brm(
  scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit"),
  chains = 4, iter = 4000, warmup = 500, cores = 8,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

scievidenc_grouped_npo_model <- brm(
  sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  save_pars = save_pars(all = TRUE),
  chains = 4, iter = 4000, cores = 8
)
sciexplain_grouped_model <- brm(
  sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit"),
  chains = 4, iter = 4000, warmup = 500, cores = 8,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

sciexplain_grouped_npo_model <- brm(
  sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  save_pars = save_pars(all = TRUE),
  chains = 4, iter = 4000, cores = 8
)
```
store(iter=4000, npo model)
```{r}
models <- list(
  ###
  ###
  ###grouped
  ##scitools_grouped_model
  scitools_grouped_model <- brm(
    scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),

  scitools_grouped_npo_model <- brm(
    scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  #scievidence_grouped_model
  scievidence_grouped_model <- brm(
    scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),

  scievidence_grouped_npo_model <- brm(
    scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  #sciexplain_grouped_model
  sciexplain_grouped_model <- brm(
    sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  sciexplain_grouped_npo_model <- brm(
    sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  
  ##ungroup
  ##ungroup
  ##ungroup
  scitools_model <- brm(
    scitools ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  scitools_npo_model <- brm(
    scitools ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  
  scievidence_model <- brm(
    scievidence ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  scievidence_npo_model <- brm(
    scievidence ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  sciexplain_model <- brm(
    sciexplain ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  sciexplain_npo_model <- brm(
    sciexplain ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  )
)

for (model_name in names(models)) {
  saveRDS(models[[model_name]], file = paste0(model_name, ".rds"))
}
```

store1
```{r}
models <- list(
  ###
  ###
  ###grouped
  ##scitools_grouped_model
  scitools_grouped_model <- brm(
    scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", disc = ~ 1),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),

  scitools_grouped_npo_model <- brm(
    scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  #scievidence_grouped_model
  scievidence_grouped_model <- brm(
    scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),

  scievidence_grouped_npo_model <- brm(
    scievidence_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  #sciexplain_grouped_model
  sciexplain_grouped_model <- brm(
    sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  sciexplain_grouped_npo_model <- brm(
    sciexplain_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  
  ##ungroup
  ##ungroup
  ##ungroup
  scitools_model <- brm(
    scitools ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  scitools_npo_model <- brm(
    scitools ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  
  scievidence_model <- brm(
    scievidence ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  scievidence_npo_model <- brm(
    scievidence ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  ),
  sciexplain_model <- brm(
    sciexplain ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit"),
    chains = 4, iter = 4000, warmup = 500, cores = 8,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99, max_treedepth = 15)
  ),
  
  sciexplain_npo_model <- brm(
    sciexplain ~ Time * session + (1 | ID) + (1 | SCHOOL),
    data = df_long,
    family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
    control = list(adapt_delta = 0.99),
    save_pars = save_pars(all = TRUE),
    chains = 4, iter = 4000, cores = 8
  )
)

for (model_name in names(models)) {
  saveRDS(models[[model_name]], file = paste0(model_name, ".rds"))
}
```

scitools
```{r}
summary(scitools_grouped_model)
# Students show a slight improvement in science conceptions post-test (Time > 0), but the effect is weak and uncertain.
# Longer curriculum (Session = 1) does not significantly improve conceptions and might even slightly decrease them.
# The interaction (Time:Session) is nearly zero, meaning the effect of Time does not differ based on session length.
# Most variability is at the individual level (sd[ID] = 1.47), meaning students differ more than schools.

summary(scitools_grouped_npo_model)
bayes_factor(scitools_grouped_model, scitools_grouped_npo_model)

loo_compare(loo(scitools_grouped_model), loo(scitools_grouped_npo_model))

#Since the NPO model (Non-Proportional Odds) does not improve fit, you can keep the simpler PO model.
plot(conditional_effects(scitools_grouped_model, "session", categorical = TRUE))
plot(conditional_effects(scitools_grouped_model, "Time", categorical = TRUE))
plot(conditional_effects(scitools_grouped_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

# plot(conditional_effects(scitools_grouped_npo_model, "session", categorical = TRUE))
# plot(conditional_effects(scitools_grouped_npo_model, "Time", categorical = TRUE))
# plot(conditional_effects(scitools_grouped_npo_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

pp_check(scitools_grouped_model)
# pp_check(scitools_grouped_npo_model)
pp_check(scitools_grouped_model, type = "scatter_avg")
#The model slightly over-predicts middle categories and under-predicts extreme categories
pp_check(scitools_grouped_model, type = "stat_2d")
# The model predicts the mean well, suggesting reasonable overall calibration.However, small discrepancies in variance may indicate missing predictors or threshold misalignment.
# pp_check(scitools_grouped_npo_model, type = "scatter_avg")
# # The NPO model performs slightly better than the PO model at predicting middle-range scores.However, it still has difficulty fully capturing category transitions, especially at the extremes.
# pp_check(scitools_grouped_npo_model, type = "stat_2d")
# # Smaller spread in predictions than in real data. Model might underestimate variability.


```
scievidence_grouped_model
```{r}
# summary(scievidence_grouped_model)
summary(scievidence_grouped_npo_model)

bayes_factor(scievidence_grouped_model, scievidence_grouped_npo_model)
loo_compare(loo(scievidence_grouped_model), loo(scievidence_grouped_npo_model))
plot(conditional_effects(scievidence_grouped_model, "session", categorical = TRUE))
plot(conditional_effects(scievidence_grouped_model, "Time", categorical = TRUE))
plot(conditional_effects(scievidence_grouped_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

# plot(conditional_effects(scievidence_grouped_npo_model, "session", categorical = TRUE))
# plot(conditional_effects(scievidence_grouped_npo_model, "Time", categorical = TRUE))
# plot(conditional_effects(scievidence_grouped_npo_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

pp_check(scievidence_grouped_model)
# pp_check(scievidence_grouped_npo_model)
pp_check(scievidence_grouped_model, type = "scatter_avg")

pp_check(scievidence_grouped_model, type = "stat_2d")

# pp_check(scievidence_grouped_npo_model, type = "scatter_avg")
# # The NPO model performs slightly better than the PO model at predicting middle-range scores.However, it still has difficulty fully capturing category transitions, especially at the extremes.
# pp_check(scievidence_grouped_npo_model, type = "stat_2d")

```

```{r}
summary(sciexplain_grouped_model)

summary(sciexplain_grouped_npo_model)
bayes_factor(sciexplain_grouped_model, sciexplain_grouped_npo_model)
loo_compare(loo(sciexplain_grouped_model), loo(sciexplain_grouped_npo_model))

plot(conditional_effects(sciexplain_grouped_model, "session", categorical = TRUE))
plot(conditional_effects(sciexplain_grouped_model, "Time", categorical = TRUE))
plot(conditional_effects(sciexplain_grouped_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

# plot(conditional_effects(sciexplain_grouped_npo_model, "session", categorical = TRUE))
# plot(conditional_effects(sciexplain_grouped_npo_model, "Time", categorical = TRUE))
# plot(conditional_effects(sciexplain_grouped_npo_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

pp_check(sciexplain_grouped_model)
# pp_check(sciexplain_grouped_model)
pp_check(sciexplain_grouped_model, type = "scatter_avg")

pp_check(sciexplain_grouped_model, type = "stat_2d")

# pp_check(sciexplain_grouped_npo_model, type = "scatter_avg")
# # The NPO model performs slightly better than the PO model at predicting middle-range scores.However, it still has difficulty fully capturing category transitions, especially at the extremes.
# pp_check(sciexplain_grouped_npo_model, type = "stat_2d")
```



Ungrouped comparison
```{r}
# scitools_grouped_model <- brm(
#   scitools_grouped ~ Time * session + (1 | ID) + (1 | SCHOOL),
#   data = df_long,
#   family = cumulative(link = "logit"),
#   chains = 4, iter = 4000, warmup = 500, cores = 8,
#   control = list(adapt_delta = 0.99, max_treedepth = 15)
# )
# scitools_model <- brm(
#   scitools ~ Time * session + (1 | ID) + (1 | SCHOOL),
#   data = df_long,
#   family = cumulative(link = "logit"),
#   chains = 4, iter = 4000, warmup = 500, cores = 8,
#   control = list(adapt_delta = 0.99, max_treedepth = 15)
# )
pp_check(scitools_npo_model, type = "stat_2d")
pp_check(scitools_npo_model, type = "scatter_avg")


bayes_factor(scitools_grouped_model, scitools_model)
scitools_npo_model <- brm(
  scitools ~ Time * session + (1 | ID) + (1 | SCHOOL),
  data = df_long,
  family = cumulative(link = "logit", threshold = "flexible"),  # Allows category-specific effects
  control = list(adapt_delta = 0.99),
  chains = 4, iter = 4000, cores = 8
)

bayes_factor(scitools_grouped_model, scitools_model)
loo_compare(loo(scitools_model), loo(scitools_npo_model))

plot(conditional_effects(scitools_model, "session", categorical = TRUE))
plot(conditional_effects(scitools_model, "Time", categorical = TRUE))
plot(conditional_effects(scitools_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

plot(conditional_effects(scitools_npo_model, "session", categorical = TRUE))
plot(conditional_effects(scitools_npo_model, "Time", categorical = TRUE))
plot(conditional_effects(scitools_npo_model, effects = "Time", categorical = TRUE,conditions = list(session = c(0, 1))))

pp_check(scitools_model)
# pp_check(scitools_npo_model)
pp_check(scitools_model, type = "scatter_avg")

pp_check(scitools_model, type = "stat_2d")

pp_check(scitools_npo_model, type = "scatter_avg")
# The NPO model performs slightly better than the PO model at predicting middle-range scores.However, it still has difficulty fully capturing category transitions, especially at the extremes.
pp_check(scitools_npo_model, type = "stat_2d")

```


```{r}
anova(scitools_grouped_model, scitools_grouped_npo_model)

loo_compare(loo(scitools_grouped_model), loo(scitools_model))
```











